A thesis draft

### 2018-11-19 res/sections/08-Background.tex
Change part on SDN

\begin{figure}
  \centering
  \includegraphics[scale=0.2]{networkslicing}
  \caption{The three layers model of network slicing}
  \label{chap:background:img:network_slicing}
\end{figure}

@@

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. The main sake of SDN is to the packet forwarding process,
\emph{data plane}, from the routing one, the so called \emph{control plane}.
The latter component serves as the centralized network intelligence and could
consists of one or more controllers. It has some drawbacks in terms of
security, scalability and elasticity, because of the centralized architecture.

\subsection{SDN stack}
\begin{figure}[ht]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be split in three different layers, called planes, based on 
their functionalities.

@@

The \emph{Application plane} role is to set the behavior of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way.
Analyzing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

@@

\emph{SDN controllers}, are logically centralized components that
receive requests from the mentioned SDN applications and either translate them 
in a fashion that can be understood from the lower layer\todo{who?}, or collect 
data needed by the application itself. 
Due to the variety of tasks a controller has to accomplish, its logical 
subdivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
that usually provide an abstract view of the network and a set of
the controller functionalities.
In particular, the middle layer represents the controller inner-logic, which 
process the requests coming from the upper layer (aka the third layer), that 
has the specific task to communicate with the remaining network components, in 
other words the Control to Data-Plane Interface and/or the already 
mentioned southbound interface.

\subsubsection{Data plane}

In the lowest part of the architecture we can find the \emph{SDN Networking 
Devices}. These ones controls the packet forwarding and data processing 
capabilities of the network. They consist of a CDPI agent and a set of one or 
multiple traffic forwarding engines and of a zero or more traffic processing 
functions, that are, as a matter of fact, the devices logic.

\subsection{OpenFlow protocol}
The SDN strategy is often associated with the OpenFlow protocol: this 
communication protocol aims is to give access to the forwarding network 
component to the lower layer network infrastructure, made of switches and 
routers. OpenFlow enables the remote administration of networks working on 
layer 3: packets forwarding tables can be modified relying on packet matching 
rules. Following this approach it is possible to change routes periodically or 
ad-hoc, based on packet types, traffic or network load.

>>>>

\begin{figure}
  \centering
  \includegraphics[scale=0.2]{networkslicing}
  \caption{The three layers model of network
  slicing~\cite{alliance2016description}}
  \label{chap:background:img:network_slicing}
\end{figure}

@@

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing and networking technology
capable of supporting the dynamic nature of future network functions and
intelligent applications while lowering operating costs through simplified
hardware, software, and management~\cite{sezer2013we}, that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. One of the main principles of SDN architecture is to decouple the
packet forwarding process, \emph{data plane}, from the routing one, the so
called \emph{control plane}. The architecture also contemplate that network
intelligence and state are logically centralized\cite{fundation2012software}.

It has some drawbacks in terms of
security, scalability and elasticity, because of the centralized architecture.

\subsection{SDN stack}
\begin{figure}[ht]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane~\cite{fundation2013software}.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be split in three different layers, also known as planes,
based on their functionalities~\cite{fundation2012software} as explained
in~\cite{fundation2013software}.

@@

In the \emph{Application plane} lays \emph{SDN applications} that are programs
whose role is to set the behavior of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI). In order to do so they can also build an
abstracted view of the network, using controllers information, in order to make
the traffic analysis easier and to simplify the decision making process..
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way. Applications consists both of
the two defined layers, one that is responsible to manage the whole logic, and
another one dedicated to the communication part, that in particular uses the
previously defined NBIs.

@@

\emph{SDN controllers}, are logically centralized components that
receive requests from SDN applications and either translate them 
in a fashion that can be understood from the SDN Datapath, or collect 
data as statics and events needed by the application to build an abstract view
of the network. 
Due to the variety of tasks a controller has to accomplish, its logical 
subdivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
to access network data and to the set of the controller functionalities.
In particular, the middle layer represents the \emph{SDN Controller Logic},
which process the requests coming from the upper layer applications. Finally
requests are communicated to the remaining network components, exploiting the
Control to Data-Plane Interface driver.

Despite the centralized design of controllers there are no limitations on the
implementation, so it possible to create federations of controllers or organize
them into a hierarchy.

\subsubsection{Data plane}
In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
network device, which exposes the possibility to access and control to its
forwarding and processing capabilities.It is composed by a CDPI agent and a set
of one or  multiple traffic forwarding engines and of a zero or more traffic
processing functions, that are, as a matter of fact, the devices logic.
Datapath can be mapped either to a single physical device or multiple ones and
the implementation of the its logic does not preclude neither the
virtualization or the sharing of physical resources nor the interoperability
with non-SDN networking.

\subsection{OpenFlow protocol}
The SDN strategy is often associated with the OpenFlow protocol because it is
the first standard communications interface defined between the control and the
datapath in SDN architecture~\cite{fundation2013software}, making possible to
access from controller the components of infrastructure, such as switches and
routers. OpenFlow enables the remote administration of networks working on 
layer 3: packets forwarding tables can be modified relying on packet matching 
rules. Following this approach it is possible to change routes periodically or 
ad-hoc, based on packet types, traffic or network load.
